---
title: "R Basics"
output: html_notebook
---
c() is a function. It’s short for “combine” or “concatenate”. It will take a sequence of comma-separated things inside the parentheses and join them together into a vector where each element is still individually accessible.

The assignment operator is <-. Think of assignment as the verb “gets”, reading left to right. Shortcut: Alt+minus

ls() displays the names of objects in your workspace

rm() removes, permanently, one or more objects from the workspace

```{r}
c("My", "name", "is", "AK")
c(TRUE, TRUE, FALSE, TRUE)
my_numbers<-c(1,2,3,1,3,5,25)
my_numbers2 <- log(my_numbers+1)
my_numbers
mean(my_numbers)
median(my_numbers)
summary(my_numbers)
sd(my_numbers)#standard deviation
var(my_numbers)#variance
cor(my_numbers, my_numbers2)#correlation
cov(my_numbers, my_numbers2)#covariance
x <- c(0, 1, 1, 2, 3, NA)#ignore NA values
sd(x, na.rm = TRUE)
#simple operations
my_numTimes5 <- my_numbers*5
rm(my_numTimes5)
my_numbers+1
my_numPower3 <- my_numbers*my_numbers*my_numbers
my_numPower3
#rm(list = ls())
```
Sequences
seq()
rep()
```{r}
1:10
seq(from = 0, to = 20)
seq(from = 0, to = 20, by = 5)
seq(from = 0, to = 20, length.out = 5)#Vector with length 5 so increments accordingly
rep(5, times = 5)
```
Comparison operators and Boolean values
```{r}
v <- c(3, pi, 4)
w <- c(pi, pi, pi)
v == w # Compare two 3-element vectors
v != w
v < w
v <= w
v > w
v >= w
any(v == pi) # Return TRUE if any element of v equals pi
all(v == 0) # Return TRUE if all elements of v are zero
rm(list = ls())
```
Indexing
First element starts at 1 and not 0
```{r}
fib <- c(0, 1, 1, 2, 3, 5, 8, 13, 21, 34)
fib[1]
fib[2]
fib[1:3]
fib[c(1, 2, 4, 8)]#Select elements anywhere within the data vector
fib[-1] # Ignore first element
fib < 10 # This vector is TRUE wherever fib is less than 10
fib[fib < 10] # Use that vector to select elements less than 10
v <- c(3, 6, 1, 9, 11, 16, 0, 3, 1, 45, 2, 8, 9, 6, -4)
v[ (v < quantile(v, 0.05)) | (v > quantile(v, 0.95)) ]#All elements in the lower and upper 5%
v <- c(1, 2, 3, 4, 5)
v[v%%2==0]#Select all elements that are even
v <- c(1, 2, 3, NA, 5)
v[!is.na(v) & !is.null(v)]#Select all elements that are neither NA nor NULL
years <- c(1960, 1964, 1976, 1994)
names(years) <- c("Kennedy", "Johnson", "Carter", "Clinton")
years
rm(list = ls())
```
Reading csv
```{r}
library(tidyverse)
url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
organs <- read_csv(file = url)
rm(list = ls())
```
Plotting Data