---
title: "R Basics"
output: html_notebook
---
c() is a function. It’s short for “combine” or “concatenate”. It will take a sequence of comma-separated things inside the parentheses and join them together into a vector where each element is still individually accessible.

The assignment operator is <-. Think of assignment as the verb “gets”, reading left to right. Shortcut: Alt+minus

ls() displays the names of objects in your workspace

rm() removes, permanently, one or more objects from the workspace

```{r}
c(1,2,5.3,6,-2,4)#numeric vector
c("My", "name", "is", "AK")#characte vector
c(TRUE, TRUE, FALSE, TRUE)#logical vector
my_numbers<-c(1,2,3,1,3,5,25)
my_numbers2 <- log(my_numbers+1)
my_numbers
mean(my_numbers)
median(my_numbers)
summary(my_numbers)
sd(my_numbers)#standard deviation
var(my_numbers)#variance
cor(my_numbers, my_numbers2)#correlation
cov(my_numbers, my_numbers2)#covariance
x <- c(0, 1, 1, 2, 3, NA)#ignore NA values
sd(x, na.rm = TRUE)
#simple operations
my_numTimes5 <- my_numbers*5
rm(my_numTimes5)
my_numbers+1
my_numPower3 <- my_numbers*my_numbers*my_numbers
my_numPower3
#rm(list = ls())
```
Sequences
seq(): function for sequences with an increment other than 1
rep(): function to create a series of repeated value
```{r}
1:10
seq(from = 0, to = 20)
seq(from = 0, to = 20, by = 5)
seq(from = 0, to = 20, length.out = 5)#Vector with length 5 so increments accordingly
rep(5, times = 5)
i <- matrix(1:20, nrow = 5, ncol = 4)
```
Comparison operators and Boolean values
```{r}
v <- c(3, pi, 4)
w <- c(pi, pi, pi)
v == w # Compare two 3-element vectors
v != w
v < w
v <= w
v > w
v >= w
any(v == pi) # Return TRUE if any element of v equals pi
all(v == 0) # Return TRUE if all elements of v are zero
rm(list = ls())
```
Indexing
First element starts at 1 and not 0
```{r}
fib <- c(0, 1, 1, 2, 3, 5, 8, 13, 21, 34)
fib[1]
fib[c(1)]
fib[2]
fib[1:3]
fib[c(1, 2, 4, 8)]#Select elements anywhere within the data vector
fib[-1] # Ignore first element
fib < 10 # This vector is TRUE wherever fib is less than 10
fib[fib < 10] # Use that vector to select elements less than 10
v <- c(3, 6, 1, 9, 11, 16, 0, 3, 1, 45, 2, 8, 9, 6, -4)
v[ (v < quantile(v, 0.05)) | (v > quantile(v, 0.95)) ]#All elements in the lower and upper 5%
v <- c(1, 2, 3, 4, 5)
v[v%%2==0]#Select all elements that are even
v <- c(1, 2, 3, NA, 5)
v[!is.na(v) & !is.null(v)]#Select all elements that are neither NA nor NULL
years <- c(1960, 1964, 1976, 1994)
names(years) <- c("Kennedy", "Johnson", "Carter", "Clinton")
years
rm(list = ls())
```
Reading csv
```{r}
library(tidyverse)
url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
organs <- read_csv(file = url)
rm(list = ls())
```
Data Visualization: Simple plots
```{r}
# Dataframes - Main idea is that each value can be of differnt modes(numeric, character, factor ,.etc)
j <- c(1,2,3,4)
k <- c("red", "white", "red", NA)
l <- c(TRUE,TRUE,TRUE,FALSE)
mydata <- data.frame(j,k,l)
names(mydata) <- c("ID","Color","Passed") # assigning column names
dim(mydata)#returns dimensions of object
head(mydata, n=10)#first 10 rows
tail(mydata, n=5) #last 10 rows
#Ordering: order(mydata$category)
#x <- mydata[order(mydata$category),mydata$category2]
#MERGE: merge(dataframeA,dataframeB,by="something"), cbind(), rbind()
library(readr)
automobiles <- read_csv("C:/Users/akshi/Desktop/automobiles.csv")
# Structure of the object/dataset
str(automobiles)
# Length of the object/dataset
length(automobiles)
# Class of the object/dataset - High level
class(automobiles)
# Type of the object/dataset - Low-level
typeof(automobiles)
# Name of the object/dataset
names(automobiles)
# Histogram plot - Used to see the distribution (Continuous data)
#hist()
# Box plot - To show the distribution of data (Min, First Quartile, Median, Third Quartile, Max)
#boxplot()
# Bar Plot - Used to compare variables (For categorical data)
#table() #need to make table of data first
#barplot(table())
# Mosaic Plot - Represents a contingency table (Distribution of two variables against each other) 
# [Box size represents the cell frequencies]
#mosaicplot()
#Customize features of graphs by modifying parameters
#par()
```
Data Visualization: ggplot()
**Build plots layer by layer**
1. Data and Mapping
2. Choose geom
3. Add layers on geom as needed
4. Organize the display aspects of the data
Basic:
ggplot(data = ..., mapping = aes (x = ..., y...))
data tells ggplot where to find the variables to use
thelibrary::function to grab a function without loading in the library
```{r}
library(gapminder)
#1
p <- ggplot(data = gapminder,mapping = aes(x = gdpPercap,y = lifeExp))
#2
p+geom_point()#scatterplot
#3
p+geom_smooth()#calculated a smoothed line for us and shaded in a ribbon showing the standard error for the line
p+geom_point()+geom_smooth(method = "lm")#lm stands for linear model
p+geom_point()+geom_smooth(method = "gam")+scale_x_log10()#scale_x_log10() scales the x-axis of a plot to a log 10 basis
p + geom_point(alpha = 0.3) + geom_smooth(method = "gam") + scale_x_log10(labels = scales::dollar) + labs(x = "GDP Per Capita", y = "Life Expectancy in Years",title = "Economic Growth and Life Expectancy",subtitle = "Data points are country-years",caption = "Source: Gapminder.")
```